---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data

Loading the data
```{r}
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
First, group the number of steps taken by date.
```{r}
total.steps <- tapply(data$steps, data$date, sum, na.rm = TRUE, simplify = TRUE)
```
Create histogram of the total number of steps taken each day
```{r}
hist(total.steps, xlab = "number of steps", main = "Histogram of the total number of steps taken each day")
```

Calculate the mean and median total number of steps taken per day.
```{r}
mean <- mean(total.steps)
print(mean)
```
```{r}
median <- median(total.steps)
print(median)
```
The mean and median total number of steps taken per day are `r mean` and `r median`

## What is the average daily activity pattern?

Calculate the average number of of steps taken in each 5-minute interval across all days.
```{r}
data$interval <- as.factor(data$interval)
step_avg <- tapply(data$steps, data$interval, sum, na.rm = TRUE, 
    simplify = TRUE)/length(levels(data$date))
```
Make a plot(time series) of the average number of steps taken.
```{r}
plot(x = levels(data$interval), y = step_avg, type = "l", xlab = "time", 
    ylab = "number of steps", main = "Average steps in 5-minute interval across all days")
```

Finding 5-minute interval containing maximum number of steps on average across all the days.
```{r}
max(step_avg)
names(which.max(step_avg))
```
```{r echo=FALSE}
intmax <- names(which.max(step_avg))
maxsteps <- max(step_avg)
```
The maximum number of steps on average across all the days is `r maxsteps` in the interval number `r intmax`.

## Imputing missing values
Calculate the total number of missing values in the dataset.
```{r}
na_num <- sum(!complete.cases(activity$steps))
```
The total number of missing values in the dataset is `r na_num`.

Use the mean/median of a 5-minute interval that has NAs to fill in all of the missing values in that interval. 

```{r}
data1 <- activity
for (i in 1:length(activity$steps)) {
    if (is.na(activity$steps[i])) {
        for (j in 1:length(step_average)) {
            if (as.character(activity$interval[i]) == names(step_average[j])) 
                activity$steps[i] = step_average[j]
        }
    }
}
data2 <- activity
clean_step_sum <- tapply(data2$steps, data2$date, sum, na.rm = TRUE, 
    simplify = TRUE)
```

```{r}
hist(clean_step_sum, xlab = "number of steps", main = "Histogram of the total number of steps taken each day")
```

## Are there differences in activity patterns between weekdays and weekends?
